\hl{Generative adversarial neural networks model of photonic crystal fiber based Surface Plasmon Resonance Sensor }

### 2020-07-24 docs/paper/paper.tex
Incoming changes from HA

\begin{abstract}
Photonic crystal fibers (PCF) for specific applications are designed and optimized by both industry experts and researches. However, the potential number of combinations possible for a single application is very large. This issue combined by the speed of the commonly used Full Vectorial Finite Element Method (FV-FEM) simulators causes the task to take significant amount of time. As stated in the previous works, artificial neural networks (ANN) can predict the result of numerical simulations much faster. However, there are two issues with the methods proposed previously: the required number of samples for training and the generality of these methods. In this paper, we propose the use of generative adversarial networks (GAN) to augment the real dataset to train an ANN model. Experimental analysis suggest that the proposed combination not only accurately predicts the confinement loss even with limited amount of data but also GAN can be used to improve existing methods in the literature. Finally, it is shown that this system can predict the confinement loss over a range of analytes and wavelengths in a completely new geometric configuration and generic enough not to require any additional tuning when used in a new dataset.
\end{abstract}

@@

In the last decade, PCF based structures extensively employed as sensing devices. Possibility of filling the cladding holes with gas and liquid analytes makes them favourable in both chemical and biological sensing applications \cite{wang2011selectively,rifat2016highly,caucheteur2015review,dash2014spr}. Moreover, due to adaptable light guidance with possibilities of controlling the confinement loss and the chromatic dispersion at any operating wavelength are distinguishing features of PCF based structures in the literature. These charming propagation features can be achieved by employing, circular, cylindrical and even square and rectangular shaped \cite{wang2015simulation} air holes in the cladding and in the core region of the structures. Furthermore, the cladding pattern arrangements such as hexagonal, octagonal, decagonal square and circular \cite{wang2011selectively,bouk2004dispersion,ademgil2014highly,razzak2007chromatic}  can be engaged to achieve the desired features for preferred applications. The hole-to hole spacing (pitch) and the air hole sizes are critical parameters of tuning the aforementioned propagation characteristics.
On the other hand, surface plasmon resonance (SPR) is a unique technique that occurs with light and free electrons interaction at metallic boundaries. This interaction results with excitement of surface plasmon polarition (SPP). Relative permittivity variations of the analytes will change the characteristics of the SPP wave, where sensing is triggered. Kretschamann Raether prism geometry is the fundamental construction principle of SPR based sensors \cite{kretschmann1968radiative}. This geometry practises Snell's Total Internal Reflection rule, where still very small amount of light refracts from the metallic layer into the evanescent field.  The refracted photons are coupling with electron charge oscillations (plasmon) at the surface of metallic layer, which are known as surface plasmon waves (SPW). The maximum energy transfer from core mode to SPWs occurs with the effective index equalisation of this corresponding modes. This phenomenon is known as resonance phase matching, where losses are reaching the maximum levels. The possible changes in the analyte index will directly affect the resonance matching wavelengths. These possible variations are the main parameters for sensitivity calculations. The deep research on SPR based sensors have shown that Kretschmann Raether’s prism geometry fails satisfying the remote sensing and miniaturization requirements of critical detecting applications. One of the most common solution of overcoming these limitations are replacing the prism with the core of an optical fiber. Considering the recent technological developments, PCF are offering tremendous amount of design flexibility with the possibilities of complete control of light guidance. Hence, SPR technique can be integrated to this unique microstructured fibers for developing comprehensive multi-purpose sensor \cite{rifat2016highly,yasli2019effect}. Following the exclusive study by Hassani and Skorobogatiy \cite{hassani2006design}, number of structures have reported the merits of PCF based SPR sensors. In the past, structures that contains features such as multi-channel single analyte and multi-channel multi analyte have been investigated \cite{rifat2016highly,yasli2019effect,zhang2011wagon,yasli2019multi,azzam2016multichannel}. Aslo, the effects of various metallic layers such as Au, Ag \cite{yasli2019multi}, Ag- Graphene \cite{rifat2016highly}, Ta2TO5 \cite{otupiri2015multi}, ITO \cite{dash2014spr}, TiO2 \cite{rifat2016highly} with various design paraeters have also been investigated in detailed.
Several design parameters such as plasmonic material, thickness of the layers, air hole size and positions are decisive on resonance matching wavelength of the SPR based structures. Optimizing and refinement of these parameters are extremely time consuming process with considerable computational time \cite{fornarelli2009neural,abdelaziz2012photonic,paper0}. In the past, genetic algorithms \cite{abdelaziz2012photonic,borel2004topology} method was used to optimize the structural design parameters to achieve better propagation features of photonic devices\cite{rifat2016highly,caucheteur2015review}. Nowadays, in order to optimize such structures, artificial neural networks (ANN) and other machine learning (ML) algorithms are being applied \cite{fornarelli2009neural,paper0,hameed2008accurate}. Although ANNs and ML have been applied to PCFs \cite{fornarelli2009neural,paper0} and SPR structures \cite{fu2018optimization,mcatee2019artificial} independently, to the best of our knowledge, there seems to have been no published papers that investigated the PCF based SPR structures with this well tested methods.

Machine Learning (ML) techniques are becoming a common tool in many fields and surpassing human performance in many tasks, such as automatic speech recognition, image recognition, natural language processing, drug discovery and toxicology. Additionally, ANN can approximate any function proven by the Universality theorem  \cite{HORNIK1991251}. This fact propelled researchers to widen the applications of ANN even further, including the study of nanophotonic structures  \cite{kiarashinejad2020knowledge}, optimization of photonic crystal nanocavities \cite{asano2018optimization}, and more recently, computing optical properties of a PCFs  \cite{paper0} and SPR structures \cite{fu2018optimization,mcatee2019artificial}.

@@

In this work, we focus on estimating the confinement loss behaviour of multichannel PCF based SPR sensor by employing the artificial neural networks. Specifically, we have constructed our system on PCF based SPR model. However, in the experiments section we have demonstrated that the designed system is generic enough to be applied to multiple similar PCF based designs. The most important contribution of this research is the use of GAN phase, where the available data is expanded to be used in the training phase.

To our knowledge, Ref.~\cite{hameed2008accurate} was the first work proposed ANN technique on PCF structures. The results demonstrated the accuracy of the network on a specific PCF structure. On the other hand, ANN method was employed for predicting the chromatic dispersion of conventional hexagonal lattice PCF structure  \cite{rodriguez2010efficient}. In Ref. \cite{mescia2011refinement}, authors employed a neural network for the refinement and design of Erbium doped PCF. Presented results have shown that conventional methods consume large amount of time compared to the runtime of the neural network. Perhaps the most rigorous experiments about neural network use in PCF design is presented in  \cite{paper0}. In this work, a feed-forward multi-layer perceptron neural network is used to calculate multiple propagation features of a PCF design. It can be evidently seen that this proposed network can achieve both acceptable accuracy and superior speed compared to the existing numerical methods. On the other hand, genetic algorithm and ANN method have also been tested on SPR sensor structures \cite{fu2018optimization,mcatee2019artificial}. It can be seen that robust performances can be achieved with these methods.

@@

This paper is organized as follows.This paper is organized as follows. The details of GAN procedure to generate additional training samples for ANN as well as the proposed neural network architecture is presented in Section \ref{sec:prop}. The design parameters and details of PCF based SPR structure is described in section \ref{sec:pcf}. Detailed analysis of the numerical results is discussed in section \ref{sec:exp}. Finally, concluding remarks are presented in Section \ref{sec:conc}.

@@

In this section, details of the proposed method are discussed. Training of the system contains two phases: a GAN phase and regression training phase. At the beginning of the training, a GAN is trained to generate additional data by using training samples. These generated samples are filtered to ensure they are within the range of the original feature space. Original training samples and generated samples are joined to train a fully-connected feedforward multi-layer perceptron (MLP) neural network. This satisfies the regression need for estimating the confinement loss. At the end of the training, the ANN will be used to decide the confinement loss of a previously unseen set of input parameters. The illustration of the proposed architecture is presented in  Figure \ref{fig:overall}. The details of the proposed GAN and ANN architecture is discussed in the following sub-sections.

@@

In our study, GAN is employed to augment the number of samples that are used in training   \cite{goodfellow2014generative}. A GAN is composed of two neural networks: generator and discriminator. The aim of the generator is to transform fully randomized data into data that follows the distribution of the original dataset. The discriminator assesses the performance of the generator and provides feedback for training. Instead of directly training the generator, it is trained through this feedback. This paradigm avoids over-fitting the data.

The GAN can be trained by diverse metrics \cite{goodfellow2014generative, mao2017least, lucic2018gans}]. In this work Wasserstein distance metric (WGAN) is selected \cite{arjovsky2017wasserstein}.In this system, discriminator is named as critic and it measures the distance between the generated and the real data. The distinguishing feature of the WGAN compared other methods is the ability to determine a stopping criteria. In a regular GAN system, training is stopped when the generated data is deemed viable by an observer. Since GAN is frequently used in image, video or audio generation, where human is used effectively in the loop. In our case, it is not viable for a human to judge the generated data. In an ordinary GAN system, removing human in-the-loop can lead to over or under-fitting, which may degrade the performance. On the other hand, WGAN uses an adaptive stopping criteria that can easily overcome the aforementioned disadvantages. Additionally, Gradient penalty is incorporated to improve the WGAN as described in \cite{gulrajani2017improved}. This proposed enhanced WGAN system converges in a stable manner without requiring further fine tuning the hyper-parameters of the system. The flowchart of this proposed system is illustrated in Figure \ref{fig:gan}.

In this work, both the critic and the generator are fully-connected feed-forward MLP models. The details of the networks are given in Table  \ref{tbl:anndetails}. The output of the system would be input and output pairs that are used to train ANN part of the system. In our sensor structure, we have 7 input parameters and a single output parameter, making a total of 8 that will be used in the GAN phase. In Section  \ref{sec:exp}, various PCF structures are tested with different inputs, resulting a different number of system parameters. The generator is supplied with the same number of parameters. These parameters are generated using Gaussian noise and is called latent variable.

Previous works \cite{ravuri2019seeing, shmelkov2018good}, demonstrated that it is possible for random data augmentation to weaken the performance of the model. Hence, it is important to sample the generated data in way to prevent performance degradation  \cite{bhattarai2019sampling}. In our proposed system, we have included a filtering step for the generated data. This step uses a simple condition to discard the values that fall out of the desired range of both the independent variables  ($n_{analyte}, \lambda, \Lambda, d1, d2, d3, n_c$), and the confinement loss. Once the training phase is completed, the generator and the filter is used to augment the number of training samples that are available for ANN to train on.

@@

In our model, we have used mini-batch training with 48 samples as the batch size. In contrast to full-batch training, mini-batch training allows more frequent gradient calculations, improving the stability and reliability of the training  \cite{masters2018revisiting, keskar2016large}. The downside of using mini-batches are added computational complexity during training. However, it has no adverse effect on the testing phase.

We have adopted Adam \cite{kingma2014adam} as the optimizer and mean squared error (MSE) as the loss/cost function. In addition, Batch Normalization algorithm \cite{ioffe2015batch} is employed to accelerate the training, reduce the effects of initial randomized state and mitigate the problem of internal covariate shift. Also, the ANN is trained using the training samples from the original dataset and the augmented samples generated by the GAN phase.

@@

The cross section of the proposed structure is illustrated in Figure 3. As can be seen from this figure, the sensor consists of analyte channel (sensing medium), plasmonic layer (Au) and various sized cladding air holes. The cladding region is formed by nineteen various sized air holes, sited on silica background with hexagonal lattice arrangement. The diameter of the air holes are set to;  $d_1$= $0.45 \mu m$, $d_2$= $0.75 \mu m$, $d_3$= $0.35 \mu m$, and  $d_c$= $0.15$. The distance between air holes ($\lambda$) are set to $2 \mu m$. Initially, trapping the light (fundamental mode) in to the core region is vital for mode quality. On the other hand, some amount of light needs to be leak towards the plasmonic layer, where the sensitivity measurement occurs. This is the key phenomenon for resonance phase matching occurrence. In this respect, the various sized cladding holes are employed for controlling the light guidance, where each set of air holes are effecting different propagation characteristic of the sensor. The Gold (Au) material is used as plasmonic layer and set to be$40 \mu m$. The Johnson and Christy Data \cite{johnson1972optical} are used the permittivity values. The Sellmeier Equation (Eq.(\ref{eqn1})) is used for calculating the refractive index of Silica material \cite{bjarklev2003PCF}.

@@

As mentioned in the introductory part, during the resonance phase matching condition, the confinement losses are reaching the highest rate. Therefore, precise calculation of the confinement losses are critical.  The Equation~\ref{eqn2} is used in this study for that purpose \cite{yasli2019effect}.

@@

A labelled dataset of 432 samples was provided by FV-FEM numerical analysis. Our dataset contains a number of independent variables, \hl{these are} the operating wavelength $\lambda$, analyte index $n_{analyte}$, hole to hole distance $\Lambda$, the air holes sizes ($d_1$, $d_2$ and $d_3$), and the real part of the refractive index ($n_c$). The labels are specified as the confinement loss of the sensor structure. The set consists of nine different configurations of the geometric properties ($\Lambda$, $d_1$, $d_2$, $d_3$), for each configuration the confinement loss was calculated for three different analytes (Water (n=1.33), Ethanol (n=1.35) and several commercial hydrocarbon mixtures (n=1.36)).

@@

In order to prove the effectiveness of the proposed setup, we have performed multiple experiments over two datasets. We have built the first dataset that is used in the experiments. Details of this dataset is explained in Section \ref{sec:pcf} and it is referred as SPR dataset. Having nine configurations, we have performed 9-fold testing on this dataset. Each fold tests a single configuration using the other configurations as training data. Slicing the data this way allowed us to guide the network to predict the loss on a configuration that has not used in training. Please note that we have applied $log_{10}$ to confinement loss to scale it.

We have used mean square error (MSE) in the comparisons. All algorithms are implemented using TensorFlow with Keras library in Python and experiments are performed on laptop with Intel i7-5600 CPU and 8GB of RAM. TensorFlow library is set to use only CPU for the experiments.

Additionally, we have used the dataset that has been used in \cite{paper0}. This dataset contains over 1000 samples. Inputs are refractive index of the core ($n_c$), wavelength ($\lambda$), hole diameter ($d$), pitch ($\Lambda$), and number of rings ($N_r$);  and outputs are effective index ($n_{eff}$), effective mode area ($A_{eff}$), dispersion ($D$), and the confinement loss ($\alpha_c$). We have only used confinement loss as the output. This dataset is referred as \dszero. Details of the datasets that are used in the experiments are given in Table \ref{tbl:dataset}. As a final note, the range and domain of these datasets are different; therefore, all numeric comparisons should be drawn within the same dataset.

@@

In this set of experiments, we have analyzed the performance of the proposed ANN design. The first experiment involves in training loss over SPR dataset. In this experiment, we have trained ANN model for up to 3000 epochs and reported the training loss in Figure \ref{chart:trainingmse}. According to this experiment, training our ANN architecture more than 2500 epochs does not yield to any improvement. Therefore, we have set the training epoch limit to 2500 in all subsequent experiments. However, this number depends on the dataset and at times additional training can improve the result. Additionally, the benefit of augmented samples from the GAN phase is evident even at the training as they reduce the oscillations in the training loss.

We have performed experiements comparing the proposed ANN architecture with the state-of-the-art method proposed in \cite{paper0}. Results of this experiment is demonstrated in Figure \ref{chart:perfnoaug}. In the experiment using SPR dataset, ANN models are predicting a new configuration and the number of samples available for training is lower. Therefore, they both have higher error rate compared to the \dszero{} dataset. According to Figure \ref{chart:perfnoaug}c, $n_{analyte} = 1.35$ case, the ANN method proposed in \cite{paper0} does not fit to the curve caused by the change in the wavelength. This can be caused by the fact that this method uses full-batch learning and therefore under-fits the model \cite{keskar2016large}. Regardless of this, both models have similar MSE due to the shift in the curve fit. This shift is caused because the network is never exposed to this particular geometric configuration.

@@

The final set of experiments shows the stability that GAN phase adds. In Table \ref{tbl:all}, we have listed minimum, maximum, average MSE along with the standard deviation between folds in the testing. It is clear that the GAN phase not only improves the average case but significantly improves the worst case while reducing the deviation, which is important in a live setting. Even in a dataset with enough examples to train the network (\dszero), using GAN phase reduces the worst case error by 32\%. This added stability can also be observed in Figure \ref{chart:trainingmse}, where the training loss has less fluctuations when the data augmentation is employed.

@@

In this paper, we have demonstrated that the use of generative adversarial networks can benefit neural networks that are designed to compute photonic crystal fiber optical properties. Additionally, we have also proposed and demonstrated a neural network that is both more generic and more accurate to estimate the confinement loss compared to the state-of-the-art network. The GAN phase also reduces the required samples to fully train the ANN, reducing the work necessary to employ the proposed system. The proposed combination is shown to be general enough that even though it is designed for the SPR dataset built for this research; in the experiments, it surpassed the accuracy of the method proposed in \cite{paper0} on the dataset this method is optimized on.

Another important aspect in this paper is the experiments performed on a dataset that is designed to test the ability of artificial neural networks to estimate the properties of a new configuration. For this purpose, we have built a database based on Surface Plasmon Resonance with varying geometric configurations. During the testing, a new configuration that is not used for the training is used. The proposed ANN, augmented with GAN phase is shown to estimate the new configuration to an acceptable error rate.

The most important benefit of using machine learning methods instead of simulation is the speed factor. In the experiments it is shown that the machine learning methods work few orders of magnitude faster than the simulation method. For different types of sensors, Convolutional and Recurrent neural networks can be of great advantage to discover the optimal patterns and geometrical shapes of the cladding air holes. This potential will be explored in a future work.

>>>>

\begin{abstract}
Photonic crystal fibers (PCF) for specific applications are designed and optimized by both industry experts and researches. However, the potential number of design combinations and flexible propagation features are offering a wide range of application areas. This issue combined by the speed of the commonly used Full Vectorial Finite Element Method (FV-FEM) simulators causes the task to take significant amount of time. As stated in the previous works, artificial neural networks (ANN) can be employed to predict the result of numerical simulations much faster. However, there are two issues with the methods proposed previously: the required number of samples for training and the generality of these methods. In this paper, we propose the use of generative adversarial networks (GAN) to augment the real dataset to train an ANN model. The experimental analysis suggested that the proposed combination not only accurately predicts the confinement loss even with limited amount of data but also GAN can be used to improve existing methods in the literature. Moreover, proposed system can also predict the confinement loss over a range of analytes and wavelengths in a completely new geometric configuration and generic enough not to require any additional tuning when used in a new dataset.
\end{abstract}

@@

In the last decade, PCF based structures extensively employed as sensing devices. Possibility of filling the cladding holes with gas and liquid analytes makes them favorable in both chemical and biological sensing applications  \cite{wang2011selectively,rifat2016highly,caucheteur2015review,dash2014spr}. Moreover, due to adaptable light guidance with possibilities of controlling the confinement loss and the chromatic dispersion at any operating wavelength are distinguishing features of PCF based structures in the literature. These charming propagation features can be achieved by employing, circular, cylindrical and even square and rectangular shaped \cite{wang2015simulation} air holes in the cladding and in the core region of the structures. Furthermore, the cladding pattern arrangements such as hexagonal, octagonal, decagonal, square and circular \cite{wang2011selectively,bouk2004dispersion,ademgil2014highly,razzak2007chromatic} can be engaged to achieve the desired features for preferred applications. The hole-to hole spacing (pitch) and the air hole sizes are critical parameters of tuning the aforementioned propagation characteristics. On the other hand, surface plasmon resonance (SPR) is a unique technique that occurs with light and free electrons interaction at metallic boundaries. This interaction results with excitement of surface plasmon polarition (SPP). Relative permittivity variations of the analytes will change the characteristics of the SPP wave, where sensing is triggered. Kretschamann Raether prism geometry is the fundamental construction principle of SPR based sensors  \cite{kretschmann1968radiative}. This geometry practices Snell’s Total Internal Reflection rule, where still very small amount of light refracts from the metallic layer into the evanescent field. The refracted photons are coupling with electron charge oscillations (plasmon) at the surface of metallic layer, which are known as surface plasmon waves (SPW). The maximum energy transfer from core mode to SPWs occurs with the effective index equalization of this corresponding modes. This phenomenon is known as resonance phase matching, where losses are reaching the maximum levels. The possible changes in the analyte index will directly affect the resonance matching wavelengths. These possible variations are the main parameters for sensitivity calculations. The deep research on SPR based sensors have shown that Kretschmann Raether’s prism geometry fails satisfying the remote sensing and miniaturization requirements of critical detecting applications. One of the most common solution of overcoming these limitations are replacing the prism with the core of an optical fiber. Considering the recent technological developments, PCF are offering tremendous amount of design flexibility with the possibilities of complete control of light guidance. Hence, SPR technique can be integrated to this unique microstructured fibers for developing comprehensive multi-purpose sensor \cite{rifat2016highly,yasli2019effect}. Following the exclusive study by Hassani and Skorobogatiy \cite{hassani2006design}, number of structures have reported the merits of PCF based SPR sensors. In the past, structures that contains features such as multi-channel single analyte and multi-channel multi analyte have been investigated \cite{rifat2016highly,yasli2019effect,zhang2011wagon,yasli2019multi,azzam2016multichannel}. Also, the effects of various metallic layers such as Au, Ag \cite{yasli2019multi}, Ag- Graphene \cite{rifat2016highly}, Ta2TO5 \cite{otupiri2015multi}, ITO \cite{dash2014spr}, TiO2 \cite{rifat2016highly} with various design parameters have also been investigated in detailed. Several design parameters such as plasmonic material, thickness of the layers, air hole size and positions are decisive on resonance matching wavelength of the SPR based structures. Optimizing and refinement of these parameters are extremely time consuming process with considerable computational time \cite{fornarelli2009neural,abdelaziz2012photonic,paper0}. In the past, genetic algorithms \cite{abdelaziz2012photonic,borel2004topology} method was used to optimize the structural design parameters to achieve better propagation features of photonic devices \cite{rifat2016highly,caucheteur2015review}. Nowadays, in order to optimize such structures, artificial neural networks (ANN) and other machine learning (ML) algorithms are being applied \cite{fornarelli2009neural,paper0,hameed2008accurate}. Although ANNs and ML have been applied to PCFs \cite{fornarelli2009neural,paper0} and SPR structures \cite{fu2018optimization,mcatee2019artificial} independently, to the best of our knowledge, there seems to have been no published papers that investigated the PCF based SPR structures with this well tested methods.

Machine Learning (ML) techniques are becoming a common tool in many fields and surpassing human performance in many tasks, such as automatic speech recognition, image recognition, natural language processing, drug discovery and toxicology. Additionally, ANN can approximate any function proven by the Universality theorem  \cite{HORNIK1991251}. This fact propelled researchers to widen the applications of ANN even further, including the study of nano-photonic structures   \cite{kiarashinejad2020knowledge}, optimization of photonic crystal nano-cavities \cite{asano2018optimization}, and more recently, computing optical properties of a PCFs  \cite{paper0} and SPR structures \cite{fu2018optimization,mcatee2019artificial}.

@@

In this work, we focused on estimating the confinement loss behavior of multichannel PCF based SPR sensor by employing the ANNs. Specifically, we have constructed our system on PCF based SPR model. However, in the experiments section we have demonstrated that the designed system is generic enough to be applied to multiple similar PCF based designs. The most important contribution of this research is the use of GAN phase, where the available data is expanded to be used in the training phase.

To our knowledge, Ref.~\cite{hameed2008accurate} was the first work proposed ANN technique on PCF structures. The results demonstrated the accuracy of the network on a specific PCF structure. On the other hand, ANN method was employed for predicting the chromatic dispersion of conventional hexagonal lattice PCF structure \cite{rodriguez2010efficient}. In Ref. \cite{mescia2011refinement},authors employed a neural network for the refinement and design of Erbium doped PCF. Presented results have shown that conventional methods consume large amount of time compared to the runtime of the neural network. Perhaps the most rigorous experiments about neural network use in PCF design is presented in  \cite{paper0}.In this work, a feed-forward multi-layer perceptron neural network is used to calculate multiple propagation features of a PCF design. It can be evidently seen that this proposed network can be achieved both acceptable accuracy and superior speed compared to the existing numerical methods. On the other hand, genetic algorithm and ANN method have also been tested on SPR sensor structures \cite{fu2018optimization,mcatee2019artificial}. It can be seen that robust performances can be achieved with these methods.

@@

This paper is organized as follows.The details of GAN procedure to generate additional training samples for ANN as well as the proposed neural network architecture is presented in Section \ref{sec:prop}. The design parameters and details of PCF based SPR structure is described in section \ref{sec:pcf}. Detailed analysis of the numerical results is discussed in section \ref{sec:exp}. Finally, concluding remarks are presented in Section \ref{sec:conc}.

@@

In this section, details of the proposed method are discussed. Training of the system contains two phases: a GAN phase and regression training phase. At the beginning of the training, a GAN is trained to generate additional data by using training samples. These generated samples are filtered to ensure the range of application. Original training samples and generated samples are joined to train a fully-connected feedforward multi-layer perceptron (MLP) neural network. This satisfies the regression need for estimating the confinement loss. At the end of the training, the ANN will be used to decide the confinement loss of a previously unexperienced set of input parameters. The illustration of the proposed architecture is presented in Figure \ref{fig:overall}. The details of the proposed GAN and ANN architecture is discussed in the following sub-sections.

@@

In our study, GAN is employed to augment the number of samples that are used in training  \cite{goodfellow2014generative}. A GAN is composed of two neural networks: generator and discriminator. The aim of the generator is to transform fully randomized data into form that follows the distribution of the original dataset. The discriminator assesses the performance of the generator and provides feedback for training. Instead of using the direct training generator, it is trained through this feedback. This paradigm avoids over-fitting the data.

The GAN can be trained by diverse metrics \cite{goodfellow2014generative, mao2017least, lucic2018gans}]. In this work Wasserstein distance metric (WGAN) is employed \cite{arjovsky2017wasserstein}. In this system, discriminator is named as critic and it measures the distance between the generated and the real data. The distinguishing feature of the WGAN compared other methods is the ability to determine a stopping criteria. In a regular GAN system, training is stopped when the generated data is deemed viable by an observer. GAN is frequently used in image, video or audio generation, where human is used effectively in the loop. In our case, it is not viable for a human to judge the generated data, therefore, automating this procedure to remove the human in-the-loop can lead to over or under-fitting, which may degrade the performance. On the other hand, WGAN uses an adaptive stopping criteria that can easily overcome the aforementioned disadvantages. Additionally, Gradient penalty is incorporated to improve the WGAN as described in \cite{gulrajani2017improved}. This proposed enhanced WGAN system converges in a stable manner without requiring further fine tuning the hyper-parameters of the system. The flowchart of this proposed system is illustrated in Figure \ref{fig:gan}.

In our work, both the critic and the generator are fully connected to feed forward MLP models. The detailed parameters of the networks are given in Table  \ref{tbl:anndetails}. The output of the system would be input and output pairs that are used to train ANN part of the system. In our proposed model, we have 7 input parameters and a single output parameter, making a total of 8 that will be used in the GAN phase. In Section  \ref{sec:exp}, various PCF structures are tested with different inputs, resulting a different number of system parameters. The generator is supplied with the same number of parameters. These parameters are generated using Gaussian noise and is called latent variable.

Previous works \cite{ravuri2019seeing, shmelkov2018good}, demonstrated that it is possible for random data augmentation to weaken the performance of the model. Hence, it is important to sample the generated data in a way to prevent performance degradation  \cite{bhattarai2019sampling}. In our proposed system, we have included a filtering step for the generated data. This step uses a simple condition to discard the values that fall out of the desired range of both the independent variables  ($n_{analyte}, \lambda, \Lambda, d1, d2, d3, n_c$), and the confinement loss. Once the training phase is completed, the generator and the filter is used to augment the number of training samples that are available for ANN to train on.

@@

In our model, we have used mini-batch training with 48 samples as the batch size. In contrast to full-batch training, mini-batch training allows more frequent gradient calculations, improving the stability and reliability of the training \cite{masters2018revisiting, keskar2016large}. The weakness of using mini-batches are added computational complexity during training. However, it has no adverse effect on the testing phase.

We have adopted Adam \cite{kingma2014adam}as the optimizer and mean squared error (MSE) as the loss/cost function. In addition, Batch Normalization algorithm \cite{ioffe2015batch} is employed to accelerate the training, reduce the effects of initial randomized state and mitigate the problem of internal covariate shift. Also, the ANN is trained using the training samples from the original dataset and the augmented samples generated by the GAN phase.

@@

The cross section of the proposed structure is illustrated in Figure 3. As can be seen from this figure, the sensor consists of analyte channel (sensing medium), plasmonic layer (Au)and various sized cladding air holes. The cladding region is formed by nineteen various sized air holes, sited on silica background with hexagonal lattice arrangement. The air hole diameters are set to;  $d_1$= $0.45 \mu m$, $d_2$= $0.75 \mu m$, $d_3$= $0.35 \mu m$, and  $d_c$= $0.15$. The distance between air holes ($\lambda$) are fixed to $2 \mu m$. Initially, trapping the light (fundamental mode) in to the core region is vital for mode quality. On the other hand, some amount of light needs to be leak towards the plasmonic layer, where the sensitivity measurement occurs. This is the key phenomenon for resonance phase matching occurrence. In this respect, the various sized cladding holes are employed for controlling the light guidance, where each set of air holes are effecting different propagation characteristic of the sensor. The Gold (Au) material is used as plasmonic layer and set to be $40 \mu m$. The Johnson and Christy Data \cite{johnson1972optical} are used the permittivity values. The Sellmeier Equation (Eq.(\ref{eqn1})) is used for calculating the refractive index of Silica material \cite{bjarklev2003PCF}.

@@

As mentioned in the introductory part, during the resonance phase matching condition, the confinement losses are reaching the highest rate. Therefore, precise calculation of the confinement losses are critical. The Equation \ref{eqn2} is used in this study for that purpose \cite{yasli2019effect}.

@@

A labelled dataset of 432 samples was provided by FVFEM numerical analysis. Our dataset contains a number of independent variables, such as the operating wavelength  $\lambda$, analyte index $n_{analyte}$, hole to hole distance $\Lambda$, the air holes sizes ($d_1$, $d_2$ and $d_3$), and the real part of the refractive index ($n_c$). The labels are specified as the confinement loss of the sensor structure. The set consists of nine different configurations of the geometric properties ($\Lambda$, $d_1$, $d_2$, $d_3$), for each configuration the confinement loss was calculated for three different analytes (Water (n=1.33), Ethanol (n=1.35) and several commercial hydrocarbon mixtures (n=1.36)).

@@

In order to prove the effectiveness of the proposed setup, we have performed multiple experiments over two datasets. We have built the first dataset that is used in the experiments. Details of this dataset is explained in Section \ref{sec:pcf} and it is referred as SPR dataset. We have performed 9-fold testing on this dataset, where nine configurations are existing. Each fold tests a single configuration using the other configurations as training data. Slicing the data allows us to guide the network to predict the loss on a configuration that has not been used in training. Please note that we have applied $log_{10}$ to confinement loss for scaling.

Also, mean square error (MSE) is used for the comparisons. All algorithms are implemented using TensorFlow with Keras library in Python and experiments are performed on device with Intel i7-5600 CPU and 8GB of RAM. TensorFlow library is set to use only CPU for the experiments.

Additionally, we have used the dataset that has been used in \cite{paper0}. This dataset contains over 1000 samples. Inputs are refractive index of the core ($n_c$), wavelength ($\lambda$), hole diameter ($d$), pitch ($\Lambda$), and number of rings ($N_r$);  and outputs are effective index ($n_{eff}$), effective mode area ($A_{eff}$), dispersion ($D$), and the confinement loss ($\alpha_c$). We have only used confinement loss as the output. This dataset is referred as PCF. Details of the datasets that are used in the experiments are given in Table \ref{tbl:dataset}. As a final note, the range and domain of these datasets are different; therefore, all numeric comparisons should be drawn within the same dataset.

@@

In this set of experiments, we have analyzed the performance of the proposed ANN design. The first experiment involves in training loss over SPR dataset. In this experiment, we have trained ANN model for up to 3000 epochs and reported the training loss in Figure \ref{chart:trainingmse}. According to this experiment, training our ANN architecture more than 2500 epochs does not yield to any improvement. Therefore, the training epoch limit is set to 2500 in all subsequent experiments. However, this number depends on the dataset and additional training may improve the results. Additionally, the benefit of augmented samples from the GAN phase is evident even at the training as they reduce the oscillations in the training loss.

We have performed experiements comparing the proposed ANN architecture with the state-of-the-art method proposed in \cite{paper0}. Results of this experiment is demonstrated in Figure \ref{chart:perfnoaug}. In SPR dataset experiment, ANN models are predicting a new configuration and the number of samples available for training is lower. Therefore, they both have higher error rate compared to the PCF dataset. According to Figure \ref{chart:perfnoaug}c, $n_{analyte} = 1.35$ case, the ANN method proposed in \cite{paper0} does not fit to the curve caused by the change in the wavelength. This can be linked to the fact that this method uses full-batch learning and therefore under-fits the model \cite{keskar2016large}. Regardless of this, both models have similar MSE due to the shift in the curve fit. Since the network is never exposed to this particular geometric configuration shift in the curve fit occurs.

@@

The final set of experiments shows the stability that GAN phase adds. In Table \ref{tbl:all},we have listed minimum, maximum, average MSE along with the standard deviation between folds in the testing. It is clear that the GAN phase not only improves the average case but significantly improves the worst case while reducing the deviation, which is important in a live setting. Even in a dataset with enough examples to train the network (PCF), using GAN phase reduces the worst case error by  32\%. This added stability can also be observed in Figure \ref{chart:trainingmse}, where the training loss has less fluctuations when the data augmentation is employed.

@@

In this paper, we have demonstrated that the use of generative adversarial networks can be beneficial to neural networks that are designed to compute propagation characteristics of PCF devices. Additionally, we have also proposed and demonstrated a neural network that is both more generic and more accurate to estimate the confinement loss compared to the state-of-the-art network. The GAN phase also reduces the required samples to fully train the ANN, reducing the work necessary to employ the proposed system. The proposed combination is shown to be general enough that even though it is designed for the SPR dataset built for this research article. In our experiments, due to optimized dataset the accuracy of the method proposed in Ref. \cite{paper0} is also improved.

Another important aspect in this work is the experiments performed on a dataset that is designed to test the ability of artificial neural networks to estimate the properties of a new configuration. For this purpose, we have built a database based on Surface Plasmon Resonance with varying geometric configurations. During the testing, a new configuration that is not used for the training is employed. The proposed ANN, augmented with GAN phase is shown to estimate the new configuration to an acceptable error rate.

The most vital advantage of using machine learning methods instead of numerical simulation is the speed factor. In the experiments it is shown that the machine learning methods work few orders of magnitude faster than the simulation method. For different types of sensors, Convolutional and Recurrent neural networks can be of great advantage to discover the optimal patterns and geometrical shapes of the cladding air holes. This potential will be explored in a future work.
